Community Detection in the Setting of Generalized Random Dot Product Graphs

### 2023-01-12 slides/slides.Rmd
reference to jcgs paper

Erdos-Renyi Model (1959)

* $P_{ij} = \theta$ (not a block model)
* 1 parameter $\theta$

@@

Popularity Adjusted Block Model (Sengupta and Chen, 2017)

@@

* $p = K (K + 1) / 2$, $q = K (K - 1) / 2$;
* $U$ is an orthogonal matrix;
* $X \in \mathbb{R}^{n \times K^2}$ is a block diagonal matrix 
composed of popularity vectors with each block corresponding to a community.

@@

$$X = \begin{bmatrix}
\Lambda^{(1)} & \cdots & 0 \\
0 & \ddots & 0 \\
0 & \cdots & \Lambda^{(K)}
\end{bmatrix}$$

\vspace*{2\baselineskip}

$$U \in \mathbb{O}(K^2)$$

$$A \sim \PABM(\{\lambda^{(k \ell)}\}_K) \iff A \sim \GRDPG_{p, q}(X U)$$

**Remark 1** (orthogonality of subspaces):
If $y_i^\top$ and $y_j^\top$ are two rows of $X U$ corresponding 
to different communities, then $y_i^\top y_j = 0$.

**Remark 2** (non-uniqueness of the latent configuration):  
If $A \sim \GRDPG_{p, q}(Y)$, then $A \sim \GRDPG_{p, q}(Y Q)$ for any $Q$ in 
the indefinite orthogonal group with signature $p, q$.

**Remark 3**: Communities correspond to subspaces even with linear transformation $Q \in \mathbb{O}(p, q)$, but this may break the orthogonality property.

@@

>>>>

@@

Popularity Adjusted Block Model (Sengupta and Chen, 2018)

@@

* $p = K (K + 1) / 2$, $q = K (K - 1) / 2$;
* $U$ is an orthogonal matrix;
* $X \in \mathbb{R}^{n \times K^2}$ is a block diagonal matrix 
composed of popularity vectors with each block corresponding to a community. 
* $P = (X U) I_{pq} (X U)^\top$.

@@

@@

* Published in JCGS.
* Code and preprint drafts available at https://github.com/johneverettkoo/pabm-grdpg
* R package WIP at https://github.com/johneverettkoo/osc
